#Analysis 2 - Top 25% of coefficient of variation small SOM, seperated by region

Purpose:

In this analysis I am checking to see if using the top 25% of genes based on co-efficient of variation using SOM clustering.  Since the clustering may be confounded between SAM and leaf tissue being forced into same cluster, as done in analysis 1, this analysis will perform clustering seperately. See `dclcmSOM_analysis2_102814.Rmd`.  

Results:

I subsetted based if the tissue was part of the SAM or leaf.  I performed the analysis on just the SAM tissue.  The PCA1 by PCA2 figure clustered along the rim.  Not sure what that means.  When I tried to cluster them together I found that they were being disappropriatly clustering.  Not sure why?  Clusters are too similar?  When looking at the clusters, it appears as though none of the genes even have the same expression.  This is not going to work.  Do SOMs need more than just 5 groups to cluster by?


Required Libraries
```{r}
library(ggplot2)
library(reshape2)
library(kohonen)
```

Cluster visualization functions 

```{r}
#ClusterVis
clusterVis <- function(clustNum){
  
  sub_cluster <- subset(plot.data, som$unit.classif==clustNum)
  sub_data <- sub_cluster[,c(1:11)] # just the sample types
  m.data <- melt(sub_data) 
  p <- ggplot(m.data, aes(x=variable, y=value))
  p + geom_point(alpha=0.5,position="jitter", size=1) + 
    geom_boxplot(alpha=0.75, outlier.size=0) + 
    theme_bw() + 
    theme(text = element_text(size=30),
          axis.text.x = element_text(angle=90, 
                                     vjust=1)) +
    xlab("Library") +
    ylab("Scaled Gene Expression")
}
```

###clusterVis_line

Need to remove unused x-axis values. 

```{r, eval=FALSE}
clusterVis_line <- function(clustNum) {
  sub_cluster <- subset(plot.data, som$unit.classif==clustNum)
  sub_data <- sub_cluster[,c(1:11)] # just the sample types
  m.data <- melt(sub_data)
  m.data$region <- ifelse(grepl("SAM", m.data$variable, ignore.case = T), "SAM", 
                          ifelse(grepl("leaf", m.data$variable, ignore.case = T), "leaf", "other"))
  head(m.data)
  m.data <- within(m.data, lineGroup <- paste(gene,sep='.'))
  ggplot(m.data, aes(variable, value, group = lineGroup)) + 
    geom_line(alpha = .1) + 
    geom_point(alpha = .0) +
    theme_bw() +
    facet_grid(.~region) +
    theme(axis.text.x = element_text(size=20, 
                                angle=90, 
                                vjust=1))
  }

```

##clusterVis_region 

not finished. 
```{r}
clusterVis_region <- function(clustNum){
  sub_cluster <- subset(plot.data, som$unit.classif==1)
  sub_data <- sub_cluster[,c(1:11)] # just the sample types
  m.data <- melt(sub_data)
  m.data$region <- ifelse(grepl("SAM", m.data$variable, ignore.case = T), "SAM", 
                          ifelse(grepl("leaf", m.data$variable, ignore.case = T), "leaf", "other"))

  m.data$stage <- ifelse(grepl("a1", m.data$variable, ignore.case = T), "1", 
                          ifelse(grepl("b2", m.data$variable, ignore.case = T), "2", 
                            ifelse(grepl("c3", m.data$variable, ignore.case = T), "3", 
                                ifelse(grepl("d4", m.data$variable, ignore.case = T), "4",
                                  ifelse(grepl("e5", m.data$variable, ignore.case = T), "5", "other")
                                  )
                                )
                             )
                          )
  
  head(m.data)       
  p <- ggplot(m.data, aes(y=value, x=variable, fill = stage))
  p + geom_point(alpha=0.5,position="jitter", size=1) + 
    geom_boxplot(alpha=0.70, outlier.size=0) +
    scale_colour_manual(values = c("darkorchid1", "coral")) +
    theme(legend.text = element_text(
      size = 30, 
      face = "bold"), 
      text = element_text(size=40)) + 
    theme_bw() + 
    theme(text = element_text(size=30)) +
    facet_grid(.~region) 
}
```

##1. Get the co-efficient of variation

```{r}
countData <- read.csv("../data/normalized_count_file.csv")
dim(countData)
#Then sort
#it adds numbers to them to make them unique but ignore
countData1 <- countData[,order(names(countData))]
countData1 <- subset(countData1, select=c(47,1:46))
```

```{r}
#remove low count libraries (3rd.leaf.7, 2nd.SAM.4, 5th.leaf.3)
dim(countData1)
names(countData1)
countData2 <- countData1[,-c(39,32,11)]
names(countData2)
dim(countData2)
countData2$a1.leaf <- rowMeans(subset(countData2[10:11]))
countData2$a1.SAM <- rowMeans(subset(countData2[12:15]))
countData2$b2.leaf <- rowMeans(subset(countData2[24:27]))
countData2$b2.SAM <- rowMeans(subset(countData2[28:30]))
countData2$c3.leaf <- rowMeans(subset(countData2[31:36]))
countData2$c3.SAM <- rowMeans(subset(countData2[37:44]))
countData2$d4.leaf <- rowMeans(subset(countData2[16:19]))
countData2$d4.SAM <- rowMeans(subset(countData2[20:23]))
countData2$e5.leaf <- rowMeans(subset(countData2[2:5]))
countData2$e5.SAM <- rowMeans(subset(countData2[6:10]))

dim(countData2)
names(countData2)
ave <- subset(countData2[45:54])
ave$sd <- apply(ave,1,function(d)sd(d))
ave$average <- rowMeans(subset(ave[1:10]))
ave$cv <- ave$sd / ave$average
dim(ave)
names(ave)

countData <- cbind(countData, countData2[45:54])
countData <- cbind(countData, ave[,11:13])

names(countData)
```

```{r}
quantile(countData$cv)
countData[is.na(countData)] <- 0
subCountData <- subset(countData, cv > 0.61264422)
allGenes25 <- subCountData[,c(1,48:60)]
colnames(allGenes25)[1]<-"gene"
```

###Subset by region

```{r}

  m.data <- melt(allGenes25)
  m.data$region <- ifelse(grepl("SAM", m.data$variable, ignore.case = T), "SAM", 
                          ifelse(grepl("leaf", m.data$variable, ignore.case = T), "leaf", "other"))
  SAM.allGenes25<- subset(m.data, region == "SAM")
  SAM.allGenes25 <- dcast(SAM.allGenes25, gene ~ variable)
  head(SAM.allGenes25)
```

##2.pca - Small

```{r}
#write.csv(allGenes25, "../data/analysis4.top25.csv")

scale_data <- as.matrix(t(scale(t(SAM.allGenes25[c(2:6)]))))
scale_data[is.na(scale_data)] <- 0 #how did NAs get reintroduced?

#Principle Component Analysis
pca <- prcomp(scale_data, scale=TRUE) 

summary(pca) 
pca.scores <- data.frame(pca$x)

SAM.allGenes25 <- cbind(SAM.allGenes25, scale_data, pca.scores) 
head(SAM.allGenes25)
```

##Visualizing the PCA

```{r}
p <- ggplot(data.val.allGenes25, aes(PC1, PC2))
p + geom_point()
```

#Self Organizing Map - (6,6) small

```{r}
data.val <- SAM.allGenes25
som.data <- as.matrix(data.val[,c(2:6)])  #subset only the scaled gene expression values
set.seed(2)
som <- som(data=som.data, somgrid(3,2,"hexagonal")) # This is where you change the size of the map
summary(som)
```

-----------------

###Training Plot ("changes") 

```{r}
plot(som, type ="changes")
```

###Code Plot - small

```{r}
plot(som, type = "codes") 
```

###Count Plot - small

```{r}
plot(som, type = "counts")
```

###Distance Neighbour Plot - Small

```{r}
plot(som, type="dist.neighbours")
```

###Heatmaps - small

```{r}
head(som$codes)
som$data <- data.frame(som$data) #changed to dataframe to extract column names easier. 

#This is just a loop that plots the distribution of each tissue type across the map. 
for (i in 1:10){
  plot(som, type = "property", property = som$codes[,i], main=names(som$data)[i])
  print(plot)
  }
```

##Visualize by Cluster

```{r}
##Bring the datasets back together for cluster specific visualizations

plot.data <- cbind(data.val[,c(1,15:34)],som$unit.classif,som$distances)
names(plot.data)
head(plot.data)
```

```{r}
p <- ggplot(data.val, aes(PC1, PC2, colour=factor(som$unit.classif))) 
p + geom_point() + theme_bw()
```

##Visualize by cluster

```{r}
clusterVis_line(1)
clusterVis_line(2)
clusterVis_line(3)
clusterVis_line(4)
clusterVis_line(5)
clusterVis_line(6)
clusterVis_line(7)
clusterVis_line(8)
clusterVis_line(9)
clusterVis_line(10)
clusterVis_line(11)
clusterVis_line(12)
clusterVis_line(13)
clusterVis_line(14)
clusterVis_line(15)
#ect
```
